---
title: "Hands-On Tutorials for Covariate Adjustment in Randomized Trials"
subtitle: "Binary Outcomes: Standardization (G-Computation)"
author: "Josh Betz (jbetz@jhu.edu), Kelly Van Lancker (kvanlan3@jhu.edu), and Michael Rosenblum (mrosen@jhu.edu)"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
header-includes:
   - \usepackage{amsmath}
output:
  ioslides_presentation:
    widescreen: TRUE
    self_contained: TRUE
bibliography: "../../bibtex_files/covariate_adjustment.bib"
---

```{r Check-Packages, include = FALSE, echo = FALSE, message = FALSE}
parameters_report <-
  config::get(
    config = "report_parameters",
    file = file.path(here::here(), "config.yml")
  )

parameters_example <-
  config::get(
    config = "binary_example_mistie",
    file = file.path(here::here(), "config.yml")
  )

installed_packages <- installed.packages()[, "Package"]

required_packages <-
  with(
    data = parameters_report,
    c(cran_packages,
      gsub(
        pattern = "^[A-Za-z0-9\\_\\.\\-]*/",
        replacement = "",
        x = github_packages
      )
    )
  )

missing_packages <-
  setdiff(
    x = required_packages,
    y = installed_packages
  )

if(length(missing_packages) > 0){
  stop(
    "Required packages not installed: ", 
    paste(missing_packages, collapse = ","), ". ",
    "See `cran_packages` and `github_packages` in the configuration file."
  )
}

example_output_path <-
  file.path(here::here(), parameters_example$example_output_file_name)

if(!file.exists(example_output_path)){
  stop(
    "Example data does not exist."
  )
} else {
  temp_environment <- new.env()
  
  load(
    file = example_output_path,
    envir = temp_environment
  )
  
  rename_params <-
    c("parameters_report")
  
  import_objects <-
    setdiff(
      x = names(temp_environment),
      y = rename_params
    )
  
  for(i in 1:length(import_objects)){
    assign(
      x = import_objects[i],
      value = get(x = import_objects[i], envir = temp_environment)
    )
  }
  
  rm(temp_environment)
}
```




```{r Report-Setup, include = FALSE, echo = FALSE, message = FALSE}
library(cobalt)
library(dplyr)
library(ggplot2)
library(ggsci)
library(jfbr)
library(kableExtra)
library(pwr)
library(rpact)
library(xfun)




### Graphic Parameters #########################################################
color_palette_1 <-
  ggsci::scale_color_lancet()
fill_palette_1 <-
  ggsci::scale_fill_lancet()

color_palette_2 <-
  ggsci::scale_color_nejm()
fill_palette_2 <-
  ggsci::scale_fill_nejm()

color_palette_3 <-
  ggsci::scale_color_bmj()
fill_palette_3 <-
  ggsci::scale_fill_bmj()

color_palette_4 <-
  ggsci::scale_color_aaas()
fill_palette_4 <-
  ggsci::scale_fill_aaas()

with(
  data = parameters_report$ggplot_parameters,
  expr = {
    assign(
      x = "plot_theme",
      value =
        theme_bw() +
      theme(
        axis.text.x = 
          element_text(
            size = axis_text_size,
            face = axis_text_face
          ),
        axis.text.y = 
          element_text(
            size = axis_text_size,
            face = axis_text_face
          ),
        axis.title = 
          element_text(
            size = axis_title_size,
            face = axis_title_face,
          ),
        strip.text =
          element_text(
            size = facet_text_size,
            face = facet_text_face,
          ),
        plot.title =
          element_text(
            size = plot_title_size,
            face = plot_title_face
          ),
        legend.text =
          element_text(
            size = legend_text_size,
            face = legend_text_face
          ),
        legend.title =
          element_text(
            size = legend_title_size,
            face = legend_title_face,
          ),
        legend.position = legend_position
      ),
      pos = .GlobalEnv
    )
  }
)


fig_w <- parameters_report$opts_chunk$fig_width
fig_h <- parameters_report$opts_chunk$fig_height

### Chunk Parameters ###########################################################
with(
  data = 
    parameters_report$opts_chunk,
  expr = {
    knitr::opts_chunk$set(
      collapse = collapse,
      echo = echo,
      message = message,
      error = error,
      purl = purl,
      results = results,
      fig.path = fig_path,
      fig.width = fig_width,
      fig.height = fig_height,
      fig.align = fig_align,
      out.width = out_width,
      dpi = dpi
    )
  }
)
```




```{r Make-Software-Table}
software_pkgs_raw <-
  dplyr::tribble(
    ~pkg,             ~c,    ~b,    ~o,  ~tte,     ~r, ~url,
    "adjrct",      FALSE, FALSE,  TRUE,  TRUE,   "GH",
    "https://github.com/nt-williams/adjrct",
    "AIPW",         TRUE,  TRUE, FALSE,  TRUE, "CRAN",
    "https://cran.r-project.org/web/packages/AIPW/index.html",
    "drord",       FALSE, FALSE,  TRUE, FALSE,   "GH",
    "https://benkeser.github.io/drord/",
    "PSW",          TRUE,  TRUE, FALSE, FALSE, "CRAN",
    "https://cran.r-project.org/web/packages/PSW/index.html",
    "RobinCar",     TRUE,  TRUE, FALSE,  TRUE, "CRAN",
    "https://cran.r-project.org/web/packages/RobinCar/index.html",
    "speff2trial",  TRUE,  TRUE, FALSE,  TRUE, "CRAN",
    "https://cran.r-project.org/web/packages/speff2trial/index.html",
    "survtmle",    FALSE, FALSE, FALSE,  TRUE,   "GH",
    "https://github.com/benkeser/survtmle",
  ) %>% 
  dplyr::rename(
    `Package` = pkg,
    `Continuous` = c,
    `Binary` = b,
    `Ordinal` = o,
    `Time-to-Event` = tte,
    `Repo` = r
  )

software_pkgs <-
  software_pkgs_raw %>% 
  dplyr::mutate(
    Package = paste0(
      "[", Package, "](", url, ")"
    ),
    dplyr::across(
      .cols = 
        dplyr::all_of(
          x = c("Continuous", "Binary", "Ordinal", "Time-to-Event")
        ),
      .fns = function(x)
        factor(
          x = x,
          levels = c(FALSE, TRUE),
          labels = c("&#9744;", "&#9745;")
        ) %>% 
        as.character
    )
  ) %>% 
  dplyr::select(
    -dplyr::any_of(x = c("url"))
  )
```

<style>
.forceBreak { -webkit-column-break-after: always; break-after: column; }
</style>


## Outline:

  - Covariate Adjustment Tutorials page
  - Worked Example: Binary Outcome - Standardization Estimator
    - Example Data: MISTIE III trial - Hemorrhagic Stroke
    - Outcome: Modified Rankin Score - Dichotomized
    - Unadjusted & Covariate Adjusted Analyses
    - Standard Errors, Confidence Intervals, p-values




# Hands-On Tutorials for Covariate Adjustment | [https://bit.ly/rct_tutorials](https://bit.ly/rct_tutorials)




## Tutorials on Covariate Adjustment | [https://bit.ly/rct_tutorials](https://bit.ly/rct_tutorials)

  - Simulated data: mimic features of trials - Scale, correlations, missingness
    - Continuous, Ordinal, Binary, Time-to-Event outcomes: Covariates
    - Example datasets: Substance Abuse, Hemorrhagic Stroke
    - Stratified randomization: improve precision
  - Common estimands of interest; Analytic approaches
  - R code for tabulating, plotting, analyzing data:
  - Compare Unadjusted & Adjusted: Benefit
  - Links to resources on learning and using R




## R Packages: Covariate Adjustment

```{r Print-Software-for-Covariate-Adjustment}
software_pkgs %>% 
  kableExtra::kbl(
    format = "html",
    escape = FALSE
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover")
  )
```




## Following Along in Code?

  - Make sure [Rtools](https://cran.r-project.org/bin/windows/Rtools/) is installed
  - Download all files from our [ENAR 2025 repository](https://github.com/CovariateAdjustment/ENAR2025)
  - Run `0_Install_Required_Packages.R` to install packages from R, GitHub
  - Run `ENAR_2025_Covariate_Adjustment_Binary.Rmd` one chunk at a time




# Example: based on MISTIE III | Functional Outcome & Mortality in Hemorrhagic Stroke




## MISTIE-III Trial: [@Hanley2019] {.smaller}

  - Hemorrhagic Stroke: Greater morbidity, mortality than ischemic stroke
    - Intracerebral Hemorrhage (ICH), possibly with Intraventricular Hemorrhage (IVH)
  - 1:1 randomized - minimally invasive surgery + thrombolytic vs. SOC medical management
  - Safety & Efficacy: Functional outcome on Modified Rankin Scale (MRS)
    - MRS at 30, 180, and 365 days post randomization
  - $Y$: Good Outcome: MRS 0-3 vs. 4-6 - independent vs. not
  - $A$: Treatment assignment: 1 = Treatment, 0 = Control
  - Design: Detect risk difference $\ge$ `r risk_difference_design` assuming $Pr\{Y|A = 0\}$ = `r p0_design`
  - Simulated data based on actual trial data: **not actual study data.**




## Simulated MISTIE Data: {.columns-2 .smaller}

  - [Codebook on covariateadjustment.github.io](https://covariateadjustment.github.io/index.html#mistie_iii)
  - Baseline Covariates
    - `age`: Age in years
    - `male`: male sex
    - `hx_cvd`:	cardiovascular disease history
    - `hx_hyperlipidemia`:	hyperlipidemia
    - `on_anticoagulants`:	on anticoagulants
    - `on_antiplatelets`:	on antiplatelets
    - `ich_location`: ICH: (`Lobar`, `Deep`)
    - `ich_s_volume`:	ICH volume on stability scan
    - `ivh_s_volume`:	IVH volume on stability scan
    - `gcs_category`: presenting Glasgow Coma Score


<p class="forceBreak"></p>


  - Treatment:
    - treatment arm: binary (`tx`) categorical (`arm`)
    - `ich_eot_volume`: intracerebral hemorrhage volume on end-of-treatment scan
  - Outcomes:
    - Modified Rankin: `_complete`: completely observed
    - `mrs_30d`: MRS at 30 days (`0-3`, `4`, `5`, `6`)
    - `mrs_180d`: MRS at 180 days (`0-2`, `3`, `4`, `5`, `6`)
    - `mrs_365d`: MRS at 365 days (`0-1`, `2`, `3`, `4`, `5`, `6`)
  - **Primary Outcome**: `mrs_365d_binary`: MRS 0-3
  - Survival: `days_on_study`, `died_on_study`




## Covariates: By Treatment, Aggregate {.smaller}

```{r Covariate-Summaries}
table1::table1(
  x = ~ 
    age + male + hx_cvd + hx_hyperlipidemia +
    on_anticoagulants + on_antiplatelets +
    ich_location + ich_s_volume + ivh_s_volume + 
    gcs_category |
    arm,
  data = sim_miii
)
```




## Covariate Balance {.smaller}

<div style="float: left; width: 50%;">

```{r Covariate-Balance, results = "markup"}
cobalt::bal.tab(
  x = 
    sim_miii %>% 
    dplyr::select(
      age, male, on_antiplatelets, ich_location, ich_s_volume,
      ivh_s_volume, gcs_category
    ),
  treat = sim_miii$arm,
  binary = "std",
  continuous = "std",
  s.d.denom = "pooled"
)
```

</div>

<div style="float: right; width: 50%;">

  - Assess balance: Standardized Differences
  - Most covariates relatively balanced:
    - Antiplatelet Use
    - IVH Volume on stability CT scan
    - GCS: Neurological Status
  - Imbalance favors medical
  
</div>




## Outcomes: {.smaller}

```{r Table-Outcomes}
table1::table1(
  x = ~ 
    mrs_365d_binary_factor |
    arm,
  data = sim_miii,
  render = jfbr::table1_render
)

y1_hat <- y1/n1
y0_hat <- y0/n0
```

```{r Crude-Estimates, echo = TRUE, eval = FALSE}
y1_hat <-
  with(data = subset(sim_miii, arm == "surgical"),
       mean(mrs_365d_binary, na.rm = TRUE))
y0_hat <-
  with(data = subset(sim_miii, arm == "medical"),
       mean(mrs_365d_binary, na.rm = TRUE))
y1_hat - y0_hat
y1_hat/y0_hat
(y1_hat*(1 - y0_hat))/(y0_hat*(1 - y1_hat))
```

  - Powered under assumption $Pr\{Y|A = 0\} \approx$ `r p0_design`, Observed $Pr\{Y|A = 0\} \approx$ `r round(x = y0_hat, digits = 2)`: Power?
  - Crude Estimates of Risk Difference, Relative Risk, Odds Ratio:
    - RD: `r round(x = y1_hat, digits = 3)` - `r round(x = y0_hat, digits = 3)` = `r round(x = y1_hat - y0_hat, digits = 3)`, RR: `r round(x = y1_hat, digits = 3)`/`r round(x = y0_hat, digits = 3)` = `r round(x = y1_hat/y0_hat, digits = 3)`, OR: `r round(x = (y1_hat*(1 - y0_hat))/(y0_hat*(1 - y1_hat)), digits = 3)`




## Standardization Estimator {.smaller}

  - **Fit** a regression model for the outcome:
    - Undjusted: $logit(Pr\{Y = 1 \vert A \}) = \beta_{0} + \beta_{A}A$
    - Adjusted: $logit(Pr\{Y = 1 \vert A \}) = \beta_{0} + \beta_{A}A + \beta_{1}X_{1} + \ldots \beta_{p}X_{p}$
  - **Predict** each individual's outcome using the fitted model
    - $\hat{Y}_{i} = logit^{-1}\{\hat{\beta}_{0} + \hat{\beta}_{A}A_{i} + \hat{\beta}_{1}X_{i1} + \ldots \hat{\beta}_{p}X_{ip}\}$
    - $\hat{Y}^{(1)}_{i} = logit^{-1}\{\hat{\beta}_{0} + \hat{\beta}_{A} + \hat{\beta}_{1}X_{i1} + \ldots \hat{\beta}_{p}X_{ip}\}$
    - $\hat{Y}^{(0)}_{i} = logit^{-1}\{\hat{\beta}_{0} + \hat{\beta}_{1}X_{i1} + \ldots \hat{\beta}_{p}X_{ip}\}$
  - **Average** these predictions over the sample: $\hat{\mu}^{(1)} = \frac{1}{n}\sum_{i=1}^{n}\hat{Y}^{(1)}_{i} \qquad \hat{\mu}^{(0)} = \frac{1}{n}\sum_{i=1}^{n}\hat{Y}^{(0)}_{i}$
  - **Contrast** the averaged predictions: $\hat{\theta}_{RD} = \hat{\mu}^{(1)} - \hat{\mu}^{(0)} \qquad \hat{\theta}_{RR} = \hat{\mu}^{(1)}/\hat{\mu}^{(0)}$




# Unadjusted Analysis | Logistic Regression




## Fitting Logistic Model {.smaller}

```{r Fit-Unadjusted-Model, echo = TRUE, results = "markup"}
mrs_unadjusted_logistic_glm <-
  stats::glm(
    formula = mrs_365d_binary ~ arm,
    data = sim_miii,
    family = binomial(link = "logit")
  )
summary(mrs_unadjusted_logistic_glm)
```




## Predict Outcomes {.columns-2 .smaller}

```{r Generate-Unadjusted-Predictions, echo = TRUE}
pr_outcome_unadj_control <-
  stats::predict(
    object = mrs_unadjusted_logistic_glm,
    newdata = 
      within(data = sim_miii,
             expr = {arm = "medical"}),
    type = "response"
  )

pr_outcome_unadj_treatment <-
  stats::predict(
    object = mrs_unadjusted_logistic_glm,
    newdata = 
      within(data = sim_miii,
             expr = {arm = "surgical"}),
    type = "response"
  )
```


<p class="forceBreak"></p>


```{r Show-Unadjusted-Predictions, echo = TRUE, results = "markup"}
table(pr_outcome_unadj_control)
table(pr_outcome_unadj_treatment)
```

  - Treatment is only covariate in model
  - Predictions are generated for each person assigning them to each treatment
  - Unadjusted predictions will be identical




## Average & Contrast Predictions  {.smaller}

<div style="float: left; width: 60%;">

```{r Average-Then-Contrast-Unadjusted-Predictions-1, echo = TRUE, results = "markup"}
e_y_0_unadj <- mean(pr_outcome_unadj_control)
e_y_1_unadj <- mean(pr_outcome_unadj_treatment)

# Risk Difference
e_y_1_unadj - e_y_0_unadj
# Relative Risk
e_y_1_unadj/e_y_0_unadj
# Odds Ratio
(e_y_1_unadj*(1 - e_y_0_unadj))/
  (e_y_0_unadj*(1 - e_y_1_unadj))
```

  - Compare two counterfactual worlds:
    - Entire population receives active intervention
    - Entire population receives comparator intervention

</div>

<div style="float: right; width: 40%;">

  - Probability of good outcome would be `r round(x = 100*rd_unadj, digits = 1)`% higher in population if everyone received surgical intervention than if everyone received standard medical care.
  - Probability of "good outcome" is `r round(x = 100*(rr_unadj - 1), digits = 1)`% greater on ratio scale
  - Odds of a "good outcome" is `r round(x = 100*(or_unadj - 1), digits = 1)`% higher on ratio scale
  - Odds Ratio overestimates relative risk: Outcome is not rare

</div>




## Average & Contrast Predictions {.columns-2 .smaller}

```{r Average-Then-Contrast-Unadjusted-Predictions-2, echo = TRUE, results = "markup"}
e_y_0_unadj <- mean(pr_outcome_unadj_control)
e_y_1_unadj <- mean(pr_outcome_unadj_treatment)

# Risk Difference
e_y_1_unadj - e_y_0_unadj
# Relative Risk
e_y_1_unadj/e_y_0_unadj
# Odds Ratio
(e_y_1_unadj*(1 - e_y_0_unadj))/
  (e_y_0_unadj*(1 - e_y_1_unadj))
```


<p class="forceBreak"></p>


```{r Compute-Unadjusted-Contrasts-from-Model, echo = TRUE, results = "markup"}
unadj_glm_beta <- coef(mrs_unadjusted_logistic_glm)
pr_medical <-
  plogis(unadj_glm_beta["(Intercept)"])

pr_surgical <-
  plogis(unadj_glm_beta["(Intercept)"] +
           unadj_glm_beta["armsurgical"])

pr_surgical - pr_medical # Risk Difference
pr_surgical/pr_medical # Relative Risk
exp(unadj_glm_beta["armsurgical"]) # Odds Ratio
```




## Standard Errors + Confidence Intervals {.smaller}

  - Adjust variance estimate for Number of Covariates $p$:
    - $v_{a} = (n - 1)/(n - p - 1)$ [@Tsiatis2008]
  - Nonparametric Bootstrap: [@Efron1994] - Draw boostrap replicate:
    - Impute missing covariates with mean, mode [@Benkeser2020]
    - Fit regression model on imputed data
    - Predict outcomes using fitted model: $\hat{Y}^{(0)}_{i}$, $\hat{Y}^{(1)}_{i}$
    - Average predictions: $\hat{\mu}^{(1)},\ \hat{\mu}^{(0)}$ - Compute Contrast: $\hat{\theta}$
  - Compute bootstrap CI $(\hat{\theta}_{L}, \hat{\theta}_{U})$ $\ $ Adjust for $p$: $\hat{\theta}_{CL(a)} = \hat{\theta} + \sqrt{v_{a}}\left(\hat{\theta}_{CL} - \hat{\theta}\right)$
  - Influence Function: $\hat{\theta} \pm Z_{(\alpha/2)}SE(\hat{\theta})\sqrt{v_{a}}$
  
$$
SE(\hat{\theta}_{RD}) = 
\sqrt{
  \frac{1}{n}
  \sum_{i=1}^n
    \left(
      2\left(2A_{i} - 1\right)\left(Y_{i} - \hat{Y}_{i}\right) + 
      \left(\hat{Y}^{(1)}_{i} - \hat{\mu}^{(1)}\right) - 
      \left(\hat{Y}^{(0)}_{i} - \hat{\mu}^{(0)}\right)
    \right)^2
}
$$




## Worked Example: Unadjusted {.smaller}

<div style="float: left; width: 60%;">

```{r Compute-Unadjusted-ATE, eval = FALSE, echo = TRUE}
source("g_computation.R")
source("boot_p_value.R")
library(boot)

set.seed(12345) # Set seed for Reproducibility

rd_unadjusted_result <-
  analysis_g_computation(
    data = sim_miii,
    formula = mrs_365d_binary ~ tx,
    family = binomial(link = "logit"),
    estimand = "difference", # "ratio" or "oddsratio"
    treatment_column = "tx",
    se_method = "bootstrap", # "none" "influence" "score"
    verbose = TRUE # Needed for p-value, not CI
  )

boot_p_value(
  boot_object = rd_unadjusted_result$boot_object,
  ci_method = "bca",
  null_value = 0,
  # Adjust CI width for DF in model
  var_adjust = rd_unadjusted_result$result$variance_factor
)
```

</div>

<div style="float: right; width: 40%;">

  - Syntax similar to regression
    - Which variable is treatment indicator
    - Which estimand is of interest
    - How to compute SE, CI
  - Function to compute p-value from bootstrap

</div>




## Results: Unadjusted

```{r Results-Unadjusted-ATE}
sim_miii_results_unadjusted %>% 
  kableExtra::kbl(
    format = "html",
    escape = FALSE,
    digits = 2
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover")
  )
```




# Covariate adjusted Analysis | Marginal Standardization (G-Computation)




## Unadjusted vs Adjusted Standardization Estimates

  - Identical: Crude vs. Logistic Regression vs. Unadjusted Standardization
  - This will not be true in general for adjusted analyses:
    - Regression: **conditional** - Standardization: **marginal**
  - Only change necessary for adjusted analysis: Add covariates to model
    - All other steps identical; Syntax identical




## Adding Covariates to Logistic Model {.smaller}

<div style="float: left; width: 35%;">

```{r Fit-Adjusted-Model-1, echo = TRUE, eval = FALSE, out.width = "30%"}
mrs_adjusted_logistic_glm <-
  stats::glm(
    formula = 
      mrs_365d_binary ~ arm +
      age + 
      male +
      hx_cvd +
      hx_hyperlipidemia +
      on_anticoagulants +
      on_antiplatelets +
      ich_location +
      ich_s_volume +
      ivh_s_volume + 
      gcs_category,
    data = sim_miii,
    family =
      binomial(link = "logit")
  )
```

</div>


<div style="float: right; width: 65%;">

```{r Fit-Adjusted-Model-2, echo = FALSE, eval = TRUE, results = "markup", out.width = "70%"}
mrs_adjusted_logistic_glm <-
  stats::glm(
    formula = 
      mrs_365d_binary ~ arm +
      age + 
      male +
      hx_cvd +
      hx_hyperlipidemia +
      on_anticoagulants +
      on_antiplatelets +
      ich_location +
      ich_s_volume +
      ivh_s_volume + 
      gcs_category,
    data = sim_miii,
    family = binomial(link = "logit")
  )

summary(mrs_adjusted_logistic_glm)
```

</div>




## Worked Example: Adjusted

  - Once model is fitted: other steps are identical:
    - Average fitted values
    - Compute Contrast: RD, RR, OR




## Worked Example: Adjusted {.smaller}

  - Identical syntax: Just add covariates to the `formula`

```{r Compute-Adjusted-ATE, eval = FALSE, echo = TRUE}
rd_adjusted_result <-
  analysis_g_computation(
    data = sim_miii,
    formula = mrs_365d_binary ~ tx + age + male + hx_cvd + hx_hyperlipidemia +
      on_anticoagulants + on_antiplatelets + ich_location + ich_s_volume +
      ivh_s_volume + gcs_category,
    family = binomial(link = "logit"),
    estimand = "difference", # "ratio" or "oddsratio"
    treatment_column = "tx",
    se_method = "bootstrap", # "none" "influence" "score"
    verbose = TRUE # Needed for p-value, not CI
  )

boot_p_value(
  boot_object = rd_adjusted_result$boot_object,
  ci_method = "bca",
  null_value = 0,
  # Adjust CI width for DF in model
  var_adjust = rd_adjusted_result$result$variance_factor
)
```




## Results: Adjusted

```{r Results-Adjusted-ATE}
sim_miii_results_adjusted %>% 
  kableExtra::kbl(
    format = "html",
    escape = FALSE,
    digits = 2
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover")
  )
```

  - Smaller Standard Error, Shorter CI
  - SE adjustment: `r round(x = sim_miii_g_computation_rd_adjusted_verbose$result$variance_factor, digits = 2)` for `r sim_miii_g_computation_rd_adjusted_verbose$g_computation_fit$n_covariates` covariates with `r sim_miii_g_computation_rd_adjusted_verbose$g_computation_fit$n_observations` observations




## Results: Unadjusted vs. Adjusted

```{r Results-Comparison}
sim_miii_results_compared %>% 
  kableExtra::kbl(
    format = "html",
    escape = FALSE,
    digits = 2
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover")
  )

re_change <- round(x = re_change, digits = 1)
rcv_change <- round(x = abs(rcv_change), digits = 1)
rcp_change <- round(x = rcp_change, digits = 1)
rciw_change <- round(x = rciw_change, digits = 1)
```

  - Efficiency Gain of `r re_change[1]`-`r re_change[2]`%
  - Reductions in variance of `r rcv_change[2]`-`r rcv_change[1]`%
  - Reduction in CI Width of `r rciw_change[1]`-`r rciw_change[2]`%




## Summary

  - Unadjusted: Crude, Unadjusted, Logistic Output - Same estimate
    - Not true in general for adjusted estimates
  - Regression Coefficients: **conditional**; Standardization: **marginal**
  - Using model: Do not assume model is correct
  - Asymptotically at least as efficient as unadjusted analysis
  - Software can target different estimands for different outcomes
    - Continuous, Binary, Ordinal, Time-to-Event
    - Adjusted or Unadjusted: Add covariates to Formula



## References {.smaller}