---
title: "Hands-On Tutorials for Covariate Adjustment in Randomized Trials"
subtitle: "Group Sequential & Information Monitoring Designs"
author: "Josh Betz (jbetz@jhu.edu), Kelly Van Lancker (kvanlan3@jhu.edu), and Michael Rosenblum (mrosen@jhu.edu)"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
header-includes:
   - \usepackage{amsmath}
output:
  ioslides_presentation:
    css: slides.css
    widescreen: TRUE
    self_contained: TRUE
bibliography: "../../bibtex_files/covariate_adjustment.bib"
---

```{r Check-Packages, include = FALSE, echo = FALSE, message = FALSE}
parameters_report <-
  config::get(
    config = "report_parameters",
    file = file.path(here::here(), "config.yml")
  )

parameters_example <-
  config::get(
    config = "binary_example_mistie",
    file = file.path(here::here(), "config.yml")
  )

installed_packages <- installed.packages()[, "Package"]

required_packages <-
  with(
    data = parameters_report,
    c(cran_packages,
      gsub(
        pattern = "^[A-Za-z0-9\\_\\.\\-]*/",
        replacement = "",
        x = github_packages
      )
    )
  )

missing_packages <-
  setdiff(
    x = required_packages,
    y = installed_packages
  )

if(length(missing_packages) > 0){
  stop(
    "Required packages not installed: ", 
    paste(missing_packages, collapse = ","), ". ",
    "See `cran_packages` and `github_packages` in the configuration file."
  )
}

example_output_path <-
  file.path(here::here(), parameters_example$example_output_file_name_gsd)

if(!file.exists(example_output_path)){
  stop(
    "Example data does not exist."
  )
} else {
  temp_environment <- new.env()
  
  load(
    file = example_output_path,
    envir = temp_environment
  )
  
  rename_params <-
    c("parameters_report")
  
  import_objects <-
    setdiff(
      x = names(temp_environment),
      y = rename_params
    )
  
  for(i in 1:length(import_objects)){
    assign(
      x = import_objects[i],
      value = get(x = import_objects[i], envir = temp_environment)
    )
  }
  
  rm(temp_environment)
}
```




```{r Report-Setup, include = FALSE, echo = FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
library(ggsci)
library(kableExtra)
library(pwr)
library(rpact)
library(xfun)




### Graphic Parameters #########################################################
color_palette_1 <-
  ggsci::scale_color_lancet()
fill_palette_1 <-
  ggsci::scale_fill_lancet()

color_palette_2 <-
  ggsci::scale_color_nejm()
fill_palette_2 <-
  ggsci::scale_fill_nejm()

color_palette_3 <-
  ggsci::scale_color_bmj()
fill_palette_3 <-
  ggsci::scale_fill_bmj()

color_palette_4 <-
  ggsci::scale_color_aaas()
fill_palette_4 <-
  ggsci::scale_fill_aaas()

with(
  data = parameters_report$ggplot_parameters,
  expr = {
    assign(
      x = "plot_theme",
      value =
        theme_bw() +
      theme(
        axis.text.x = 
          element_text(
            size = axis_text_size,
            face = axis_text_face
          ),
        axis.text.y = 
          element_text(
            size = axis_text_size,
            face = axis_text_face
          ),
        axis.title = 
          element_text(
            size = axis_title_size,
            face = axis_title_face,
          ),
        strip.text =
          element_text(
            size = facet_text_size,
            face = facet_text_face,
          ),
        plot.title =
          element_text(
            size = plot_title_size,
            face = plot_title_face
          ),
        legend.text =
          element_text(
            size = legend_text_size,
            face = legend_text_face
          ),
        legend.title =
          element_text(
            size = legend_title_size,
            face = legend_title_face,
          ),
        legend.position = legend_position
      ),
      pos = .GlobalEnv
    )
  }
)


fig_w <- parameters_report$opts_chunk$fig_width
fig_h <- parameters_report$opts_chunk$fig_height

### Chunk Parameters ###########################################################
with(
  data = 
    parameters_report$opts_chunk,
  expr = {
    knitr::opts_chunk$set(
      collapse = collapse,
      echo = echo,
      message = message,
      error = error,
      purl = purl,
      results = results,
      fig.path = fig_path,
      fig.width = fig_width,
      fig.height = fig_height,
      fig.align = fig_align,
      out.width = out_width,
      dpi = dpi
    )
  }
)
```




## Outline

  - Overview: Open Source Software:
    - Group Sequential Designs, Information Monitoring
  - Review Fixed Sample Size Plans in **R**
  - Convert Fixed Sample Size to Group Sequential Design
  - Analyze Example Datasets




## Software

  - Planning, Analyzing Group Sequential Designs:
    - [rpact](https://www.rpact.org): [Tutorials & Vignettes](https://www.rpact.org/#learn-to-use-rpact)
    - [gsDesign](https://keaven.github.io/gsDesign/): [Web Interface](https://keaven.github.io/gsd-shiny/) and [related software](https://keaven.github.io/software/)
  - Planning, Monitoring, Analyzing Information Monitored Designs
    - [impart](https://jbetz-jhu.github.io/impart/) - [Tutorials &  Vignettes](https://jbetz-jhu.github.io/impart/articles/)




## MISTIE-III Trial - Effect Size $h$: [@Hanley2019] {.smaller}

  - Outcome: Dichotomized Modified Rankin Scale: 
    - $0-3 \rightarrow 1$: Good Functional Outcome - Alive, Independent
    - $4-6 \rightarrow 0$: Poor Functional Outcome
  - Assumed $Pr\{Y|A = 0\} = \pi_{0} \approx$ `r p0_design`: 
  - Detect Risk Difference $\delta_{RD} \ge$ `r risk_difference_design`: $Pr\{Y|A = 1\} = \pi_{1} \ge$ `r p1_design`
  - Power = `r power_design`, Type I Error = `r alpha_design`: Two-sided test, `r 100*pr_missing`% missing data
    - Just for demonstration: Not values from original study

```{r Design-Parameters, echo = TRUE, eval = FALSE}
p0_design <- 0.25
risk_difference_design <- 0.13
p1_design <- p0_design + risk_difference_design
alpha_design <- 0.05
power_design <- 0.80
pr_missing <- 0.20
```



## MISTIE-III Trial - Fixed Sample Size $N$: {.smaller}

  - Determine Effect size $h = 2\left(sin^{-1}\left(\sqrt{\pi_{1}}\right) - sin^{-1}\left(\sqrt{\pi_{0}}\right) \right) =$ `r round(es_h_design, digits = 3)`

```{r Effect-Size-h, echo = TRUE, results = "markup"}
pwr::ES.h(p1 = p1_design, p2 = p0_design)
```

  - Determine $N$ **per arm** of complete observations

```{r N-Per-Arm-Unadjusted, echo = TRUE, results = "markup"}
n_per_arm_fixed <-
  pwr::pwr.2p.test(
    h = es_h_design,
    sig.level = alpha_design,
    power = power_design
  )$n %>% ceiling()

n_per_arm_fixed
```

  - We need `r 2*n_per_arm_fixed` complete outcomes: Divide by fraction complete = `r (1 - pr_missing)` $\rightarrow$ `r n_total_fixed` participants

```{r N-Per-Arm-Adjusted-Missing, echo = TRUE, results = "markup"}
ceiling(2*n_per_arm_fixed/(1 - pr_missing))
```




## Group Sequential Design (GSD) {.smaller}

  - 3 Stages: 2 Interim Analyses (50%, 75% of $N$) + Final Analysis
  - Efficacy Bounds: O'Brien Fleming $\alpha$-spending Functions - No Futility

```{r Set-Up-GSD, echo = TRUE, eval = FALSE}
information_rates <- c(0.50, 0.75, 1.00)
type_of_design_asof <- "asOF"
type_beta_spending_asof <- "none"

trial_design <-
  rpact::getDesignGroupSequential(
    alpha = alpha_design,
    beta = 1 - power_design,
    sided = test_sides,
    informationRates = information_rates,
    typeOfDesign = type_of_design_asof,
    typeBetaSpending = type_beta_spending_asof
  )

plot(trial_design) # Plot Boundaries
```




## O'Brien-Fleming: Efficacy Only {.smaller}

```{r Plot-GSD-ASOF-Boundaries, fig.width = 8, fig.height = 8, out.width = "50%", out.height = "50%"}
plot(trial_design_asof)
```




## O'Brien-Fleming: Efficacy & Futility {.smaller}

```{r Plot-GSD-ASOF-BSOF-Boundaries, fig.width = 8, fig.height = 8, out.width = "50%", out.height = "50%"}
plot(trial_design_asof_bsof)
```



## Compute Sample Size Adjustment

  - Up to 3 interim analyses: $\uparrow N$ to maintain Type I error $\alpha$ 

```{r Compute-Inflation-Factor, eval = FALSE, echo = TRUE}
inflation_factor <-
  rpact::getDesignCharacteristics(
    design = trial_design_asof
  )$inflationFactor
n_total_gsd_asof <-
  ceiling(2*n_per_arm_fixed*inflation_factor_asof)
n_at_analyses_asof <-
  ceiling(information_rates*n_total_gsd_asof)
```
  
  - Maximum Sample Size is `r round(x = 100*(inflation_factor_asof - 1), digits = 1)`% larger than Fixed Sample Size
  - Potential to stop at $N =$ `r n_at_analyses_asof[1]` or `r n_at_analyses_asof[2]`




# Analyzing Group Sequential Designs




## Using the `impart` Package: {.smaller}

  - **I**nformation **m**onitoring for **p**recision **a**daptive **r**andomized **t**rials
    - Using Covariate Adjustment in Group Sequential Designs: Focus
    - Designing, Monitoring, Analyzing trials with Information Monitoring: Later
  - Group Sequential Designs:
    - Set up a GSD with [rpact](https://www.rpact.org) package
    - Set up a `monitored_design`: Stores settings for analyses: $N_{max}$, RNG seeds, $\theta_{0}$
    - Perform analyses: gives an updated `monitored_design` to pass to next analysis
      - Keep going until stop for efficacy, futility, or fail to reject
    - Final analysis will contain entire history of study:
  - Users can supply their own functions for analysis:
    - Standardization (Continuous, binomial); Marginal HR (Time-to-Event)




## `impart` - Setting Up a Design {.smaller}

```{r Set-Up-Design, echo = TRUE, eval = FALSE}
# Set up monitored design - **See .Rmd for details**
monitored_design_asof <-
  initialize_monitored_design(
    # Trial Design set up using rpact
    trial_design = trial_design_asof,
    # Null value of parameter
    null_value = 0,
    # Maximum Sample Size
    maximum_sample_size = n_total_gsd_asof,
    # Target precision: Based on assumed effect, Power, Alpha
    information_target = information_adaptive_asof,
    # Enforce Independent Increments with Orthogonalization
    orthogonalize = TRUE,
    # RNG Seed for analyses
    rng_seed_analysis = 54321
  )
```




## Interim Analysis 1/3: O-F Efficacy {.smaller}

```{r Analysis-1-ASOF, echo = TRUE, eval = FALSE}
interim_analysis_asof_1 <-
  monitored_analysis(
    data = sim_miii_asof_ia_1, # Current Data
    monitored_design = monitored_design_asof, # Design Details
    estimation_function = standardization, # Function used for analysis
    estimation_arguments = # Arguments that `estimation_function` uses
      list(
        estimand = "difference",
        outcome_formula = 
          mrs_356d_binary ~ tx + age + male + hx_cvd + hx_hyperlipidemia +
          on_anticoagulants + on_antiplatelets + ich_location + ich_s_volume +
          ivh_s_volume + gcs_category,
        family = binomial,
        treatment_column = "tx",
        outcome_indicator_column = ".r_3"
      ),
    # Optional function to adjust variance (e.g. for # of covariates)
    correction_function = standardization_correction
  )
```




## Interim Analysis 1 Results: {.smaller}

```{r Analysis-1-ASOF-Result, echo = TRUE, results = 'markup'}
# Decision
interim_analysis_asof_1$interim_analysis_1$decision_detail

# Test statistics, Boundaries
interim_analysis_asof_1$interim_analysis_1$decision_data
```

  - Test statistic hasn't exceeded any efficacy boundaries




## Interim Analysis 2/3: O-F Efficacy {.smaller}

```{r Analysis-2-ASOF, echo = TRUE, eval = FALSE}
interim_analysis_asof_2 <-
  monitored_analysis(
    
    data = sim_miii_asof_ia_2, # Pass updated data
    monitored_design = interim_analysis_asof_1, # Pass previous analysis
    
    # Specify what to do for interim analysis 2
    estimation_function = standardization,
    estimation_arguments = 
      list(
        estimand = "difference",
        outcome_formula = 
          mrs_356d_binary ~ tx + age + male + hx_cvd + hx_hyperlipidemia +
          on_anticoagulants + on_antiplatelets + ich_location + ich_s_volume +
          ivh_s_volume + gcs_category,
        family = binomial,
        treatment_column = "tx",
        outcome_indicator_column = ".r_3"
      ),
    
    correction_function = standardization_correction
  )
```




## Interim Analysis 2 Results: {.smaller}

```{r Analysis-2-ASOF-Result, echo = TRUE, results = 'markup'}
# Decision
interim_analysis_asof_2$interim_analysis_2$decision_detail

# Test statistics, Boundaries
interim_analysis_asof_2$interim_analysis_2$decision_data

# Entire history of results are nested within the result
names(interim_analysis_asof_2)
```

  - Test statistic hasn't exceeded any efficacy boundaries


## Final Analysis: O-F Efficacy {.smaller}

```{r Analysis-3-ASOF, echo = TRUE, eval = FALSE}
final_analysis_asof <-
  monitored_analysis(
    
    data = sim_miii_asof_ia_3, # Pass updated data
    monitored_design = interim_analysis_asof_2, # Pass previous analysis
    
    # Specify what to do for interim analysis 2
    estimation_function = standardization,
    estimation_arguments = 
      list(
        estimand = "difference",
        outcome_formula = 
          mrs_356d_binary ~ tx + age + male + hx_cvd + hx_hyperlipidemia +
          on_anticoagulants + on_antiplatelets + ich_location + ich_s_volume +
          ivh_s_volume + gcs_category,
        family = binomial,
        treatment_column = "tx",
        outcome_indicator_column = ".r_3"
      ),
    
    correction_function = standardization_correction
  )
```




## Final Analysis Results: {.smaller}

```{r Analysis-3-ASOF-Result, echo = TRUE, results = 'markup'}
# Decision
final_analysis_asof$final_analysis$decision_detail

# Test statistics, Boundaries
final_analysis_asof$final_analysis$decision_data

# Entire history of results are nested within the result
names(final_analysis_asof)
```




# Add Futility Stopping:




## O-F Efficacy **+ Futility** - Analysis 1 {.smaller}

```{r Analysis-1-ASOF-BSOF, echo = TRUE, eval = FALSE}
interim_analysis_asof_bsof_1 <-
  monitored_analysis(
    
    data = sim_miii_asof_bsof_ia_1,
    monitored_design = monitored_design_asof_bsof, # Design Details
    
    estimation_function = standardization,
    estimation_arguments =
      list(
        estimand = "difference",
        outcome_formula = 
          mrs_356d_binary ~ tx + age + male + hx_cvd + hx_hyperlipidemia +
          on_anticoagulants + on_antiplatelets + ich_location + ich_s_volume +
          ivh_s_volume + gcs_category,
        family = binomial,
        treatment_column = "tx",
        outcome_indicator_column = ".r_3"
      ),
    correction_function = standardization_correction
  )
```




## Interim Analysis 1 Results: {.smaller}

```{r Analysis-1-ASOF-BSOF-Result, echo = TRUE, results = 'markup'}
# Decision
interim_analysis_asof_bsof_1$interim_analysis_1$decision_detail

# Test statistics, Boundaries
interim_analysis_asof_bsof_1$interim_analysis_1$decision_data
```

  - Recommend to stop: Futility





## Conclusions {.smaller}

  - [rpact](https://www.rpact.org) and [gsDesign](https://keaven.github.io/gsDesign/): Great software, vignettes - Also see [@Jennison1999]
  - Specify Initial Design:
  - First Analysis: Pass Design, Analysis + Arguments, Data
  - Additional Analyses:
    - Pass previous result, Analysis + Arguments, Updated Data
  - Repeat until stopping criteria met: Each result nested in object


## References {.smaller}